---
title: "Who does more pushups"
author: "McKayla Witt"
date: "OCT. 2017"
output: html_document
---

```{r include=FALSE}
require(knitr)
require(mosaic)
require(tigerstats)
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

Several years ago some MAT 111 students who were on the footbal team conducted a study of their team-mates.  The results are contained in the data frame `pushups` in the `tigerstats` package.

```{r eval=FALSE}
data(pushups)
View(pushups)
help(pushups)
```

Among other things, the students recorded the type of position a player had on the team (either line or skill), and also the number of push-ups each player could perform in a two-minute period.

Line players are on the offensive or defensive line.  They need to be a certain kind of speed, but they must also be quite big and strong.  Skill players do a lot of runninga and catching:  they are the quarterback, running backs, receivers, etc.

Our Research Que stion is:

  >Who does more pushups, on average: a line player or a skill player?


# Methods
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

<<<<<<< HEAD
To fully analyze the data provided by Georgetown, the graphical results will contain a scatterplot displaying the weight and amount of pushups achieved by each player. From there the scatter plot will be color-coordinated to reflect the players' position on the football team.
To centralize the data we are analyzing, specifically the average pushups completed by players based on whether they play as a "LINE" or "SKILL", I will create a box plot.
=======
```{r}
pushups_skill <- c(46, 38, 31, 23, 32, 18, 52, 27, 40, 58, 47, 108, 62, 56, 92, 50, 88, 31, 30, 68, 49, 48)
pushups_line <- c(8, 6, 32, 11, 12, 17, 24, 25)


mean(pushups_skill) 
mean(pushups_line)

average_pushups_skill <- mean(pushups_skill)
average_pushups_line <- mean(pushups_line)


average_pushups_skill > average_pushups_line
```



>>>>>>> methods

The numerical results will be re-orgainized into "pushups_skill" and "pushups_line". Then the data will be computed to find the mean of each data set. These results will be renamed to "average_pushups_skill" and "average_pushups_line" respectively. The afformentioned vectors will then be compared. 
To verify that my initial work is correct, I will run the orginal Georgetown data set through a t.test.


# Results
<!--You have both graphical and numerical methods, so you will have both graphical and numerical results to consider.!-->

## Graphical Results
<!--Your graph should have a good title, and axes should be labled as needed.  Discuss your graphical results, showing how they answer the research question.  Use specific features of the graph in your discussion.!-->

```{r}
xyplot(weight ~ pushups,
	data = pushups,
	groups = position,
	auto.key = list(
		space = "top",
		title = "position",
		cex.title = 1,
		columns = 1))
```

```{r}
bwplot(pushups ~ position,
	data = pushups)
```


## Numerical Results
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
pushups_skill <- c(46, 38, 31, 23, 32, 18, 52, 27, 40, 58, 47, 108, 62, 56, 92, 50, 88, 31, 30, 68, 49, 48)
pushups_line <- c(8, 6, 32, 11, 12, 17, 24, 25)


mean(pushups_skill) 
mean(pushups_line)

average_pushups_skill <- mean(pushups_skill)
average_pushups_line <- mean(pushups_line)


average_pushups_skill > average_pushups_line
```

```{r}
t.test(pushups~position, data=pushups)
```



# Inferential Results
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc. Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
The t.test, verified that the "SKILL" players do more pushups on average than the "LINE" players. "LINE" players group average 17.125 pushups, while "SKILL" players average 49.72727 pushups. Because the P-Value is 9.421e-06, the chances that there are sampling or experimemtal errors that would deem the difference insignificant are slim. Therefore the Null hypothesis can be dismissed. 

The assumptions this test was based on are that the Georgetown data is complete. Based on that assumption there is no reason to doubt the results of these results, unless it came to light that players had been left out of the study. 

# Discussion and Conclusion
<!--State your overall conclusion.  This is also a good place to mention other Research Questions that came to mind as you were carrying out this study.!-->
The results show that the "SKILL" players complete more pushups on average than the "LINE" players. After reaching this conclusion and throughly studying the graphs, I wonder if weight has any impact on the differences between averages. 
